A Topic Model of Climate Change Literature

### 2017-05-11 main.tex

\documentclass{article}
\usepackage[a4paper, total={7in, 8in}]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{natbib}
\usepackage{todonotes}
\usepackage{booktabs}
\title{A Topic Model of Climate Change Literature}
\title{Words, words, words: Mapping the Matter of Climate Change Literature}
\title{A Topography of Climate Change Research}
\author{Max Callaghan}
\begin{document}
\maketitle
\textbf{
To contribute to evidence-based policies on tackling climate change, the IPCC aims to comprehensively assess the relevant scientific literature \citep{IPCC2013}. With the size of this literature currently at least [x] times larger than at the time of the IPCC's first assessment report \citep{Houghton1990ClimateAssessment}, this task has become impossible without the aid of machine-reading. We collect over 300,000 abstracts from Web of Science (WoS) and Scopus, and develop a topic model in order to give an overview of this unmanageably large corpus. We find that new topics, such as [biochar] can be identified as they emerge, an application with a high potential for increasing the relevance and comprehensiveness of further IPCC reports.
}

The size of the scientific literature on climate change has expanded rapidly over the lifetime of the IPCC. While the first assessment report had around 7,000 articles to assess, 5,000 new articles are now published every month, bringing the total size of the literature to well over half a million papers, (Figure \ref{growth}). The increase in volume, velocity, and variety of content to be assessed has turned the task of the IPCC into a `Big Literature` challenge. To ask questions about the literature \textit{at scale}, we now need to employ computational techniques involving natural language processing.

@@

\begin{figure}
	\includegraphics[width=\linewidth]{plots/hot_topics}
    \caption{Topic growth over time. The 3 topics in each assessment period that grew by the largest amount are labelled}
\end{figure}

@@

\section{Introduction}
\begin{itemize}
	\item Literature exploding, IPCC not keeping up \citep{minx2016learning}
    \item No systematic way of selecting references, no comprehensive assessment
    \item First step towards this is a map provided by topic models. 
    \item A map shows the places which the assessment makers need to navigate, leaving them to decide course. Computer assisted not computer decided.
    \item What is the topic structure of climate change research? How has it changed over time?
\end{itemize}

\section{Methodology}

\section{Results}
\begin{itemize}
	\item The biggest topics are x and y
	\item These topics have grown, these have decreased
    \item X keywords fit into Y topics like so...
    \item Finding similar documents works across topics, rather than just by keywords, so better.
    \item Similar documents more or less likely to be across disciplinary boundaries ??
    \item Model selection validated by these measurements.
\end{itemize}

\section{Conclusion}
\begin{itemize}
	\item A very simple topic model provides an overview of the whole landscape.
    \item This allows researchers / assessment makers to identify areas that have grown recently
    \item Topic models aid document discovery, have the potential to contribute to more comprehensive assessments.
    \item Next steps for research: using topic models to assess the assessment process: find gaps etc.
\end{itemize}

@@

\listoffigures

\bibliography{Mendeley.bib}
\bibliographystyle{apalike}

>>>>

\documentclass{article}
\usepackage[a4paper, total={7in, 8in}]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{natbib}
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{lineno}

\linespread{2}

% Title
\title{A Topic Model of Climate Change Literature}
\title{Words, words, words: Mapping the Matter of Climate Change Literature}
\title{A Topography of Climate Change Research}
\author{Max Callaghan}

\begin{document}
\maketitle

\linenumbers

\textbf{
To contribute to evidence-based policies on tackling climate change, the IPCC aims to comprehensively assess the relevant scientific literature \citep{IPCC2013}. With the size of this literature currently at least 100 times larger than at the time of the IPCC's first assessment report \citep{Houghton1990ClimateAssessment}, this task has become impossible without the aid of machine-reading. We collect over 600,000 abstracts from Web of Science (WoS) and Scopus, and develop a topic model in order to give an overview of this unmanageably large corpus. We find that new topics, such as [biochar] can be identified as they emerge, an application with a high potential for increasing the relevance and comprehensiveness of further IPCC reports.
}

The size of the scientific literature on climate change has expanded rapidly over the lifetime of the IPCC. While the first assessment report had around 7,000 articles to assess, 5,000 new articles are now published every month, bringing the total size of the literature to well over half a million papers, (Figure \ref{growth}). The increase in volume, velocity, and variety of content to be assessed has turned the task of the IPCC into a `Big Literature' challenge. To ask questions about the literature \textit{at scale}, we now need to apply computational techniques to the analysis of large document collections.

@@

\begin{figure}
	\includegraphics[width=\linewidth]{plots/hot_topics}
    \caption{Topic growth over time. The 3 topics in each assessment period that grew by the largest amount are labelled [Example figure]}
\end{figure}

@@

\section*{Results}
\begin{itemize}
	\item The biggest topics are x and y
	\item These topics have grown, these have decreased
    \item X keywords fit into Y topics like so...
    \item Finding similar documents works across topics, rather than just by keywords, so better.
    \item Similar documents more or less likely to be across disciplinary boundaries ??
    \item Model selection validated by these measurements.
\end{itemize}

\section{*Conclusion}
\begin{itemize}
	\item A very simple topic model provides an overview of the whole landscape.
    \item This allows researchers / assessment makers to identify areas that have grown recently
    \item Topic models aid document discovery, have the potential to contribute to more comprehensive assessments.
    \item Next steps for research: using topic models to assess the assessment process: find gaps etc.
\end{itemize}

@@

\section*{Methodology}

\listoffigures
\linespread{1}
\bibliography{Mendeley.bib}
\bibliographystyle{apalike}
